---
title: "CRC_inteSIN"
author: "Huahui Ren"
date: "2024-03-15"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduce 

Here is the test about the  classification power of CRC vs Control using sample-specific inference network methods.

## Content

- Part 1: Preprocessing 
- Part 2: Comparisions in biomarkers using relative abundance or degree 
- Part 3: Comparision in classifical performance 

## Part 1: 

```{r , message=FALSE, warning=F}
library(tidyverse)

crc_data <- readRDS("data/crc_infan.rds")
Yachi_meta <- crc_data$Yachidas_meta[[2]]  %>% as.data.frame() %>% mutate(id = rownames(.))
Yuj_meta <- crc_data$yuj_meta[[2]] %>% as.data.frame() %>% mutate(id = rownames(.))
Wir_meta <- crc_data$WirbelJ_meta[[2]]  %>% as.data.frame() %>% mutate(id = rownames(.))

crc_meta <- Yachi_meta %>%
  full_join(Yuj_meta, by = "id") %>%
  full_join(Wir_meta, by = "id") %>% column_to_rownames(var = "id")
crc_meta[is.na(crc_meta)] <- 0

library(rmeta)
crc_meta_f <- filterPer(crc_meta, row = 1, percent = 0.1) |> t() |> renorm()
crc_phe <- do.call("rbind", list(crc_data$Yachidas_meta[[1]], crc_data$yuj_meta[[1]], crc_data$WirbelJ_meta[[1]]))

crc_phe_f <- crc_phe |> filter(disease == "CRC" | disease == "healthy" |disease == "CRC;T2D"|disease == "T2D") 
rownames(crc_phe_f) <- crc_phe_f$sample_id
crc_phe_f$disease <- ifelse(crc_phe_f$disease == "CRC;T2D", "CRC", ifelse(crc_phe_f$disease == "T2D", "healthy", crc_phe_f$disease))
crc_meta_f <- crc_meta_f[rownames(crc_phe_f), ]

library(gtsummary)

crc_phe_f |> select(study_name, disease, age, gender, BMI) |> tbl_summary(by = study_name) |> add_p() |>
  add_overall() |>
  bold_labels()

```

## Part 2: 

### Overview of core species in 3 crc cohort 

*1.* WirbelJ_2018

```{r, message=F, error=F, warning=F}
source("R/tinyFun.R")
library(RColorBrewer)
prevalences <- seq(0.05, 1, 0.05)
detections <- 10^seq(log10(1e-3), log10(.2), length = 10)

WirbelJ_2018_meta <- crc_meta_f[crc_phe_f$study_name == "WirbelJ_2018", ]

core_tax_plot(dat = WirbelJ_2018_meta, 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))

```

```{r}
pcoa_study1 <- ggPCoA(data = as.data.frame(WirbelJ_2018_meta), group = crc_phe_f[crc_phe_f$study_name == "WirbelJ_2018", "disease"],  color = c("#4E6E81", "#F9DBBB"))

pcoa_study1
```

## 

*2.* YachidaS_2019

```{r}
YachidaS_2019_meta <- crc_meta_f[crc_phe_f$study_name == "YachidaS_2019", ]

core_tax_plot(dat = YachidaS_2019_meta , 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))
```

```{r}
pcoa_study2 <- ggPCoA(data = as.data.frame(YachidaS_2019_meta), group = crc_phe_f[crc_phe_f$study_name == "YachidaS_2019", "disease"],  color = c("#4E6E81", "#F9DBBB"))

pcoa_study2
```

## 

*3.* YuJ_2015

```{r}
YuJ_2015_meta <- crc_meta_f[crc_phe_f$study_name == "YuJ_2015", ]

core_tax_plot(dat = YuJ_2015_meta , 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))
```

```{r}
pcoa_study3 <- ggPCoA(data = as.data.frame(YuJ_2015_meta), group = crc_phe_f[crc_phe_f$study_name == "YuJ_2015", "disease"],  color = c("#4E6E81", "#F9DBBB"))
pcoa_study3
```

```{r, fig.width=12, fig.height=4}
library(patchwork)
pcoa_study1+pcoa_study2+pcoa_study3

```

## 

### Compare crc vs healthy using relative abundance  

```{r dplyr, warning=F, message=F, fig.width=12, fig.height=5}
library(coin)
source("R/tinyFun.R")
#first try to clr or alr transformation 
crc_meta_clr <- robCompositions::cenLR(zCompositions::cmultRepl(crc_meta_f, z.warning = 0.9))[[1]]
#crc_meta_clr <- vegan::decostand(crc_meta_f, method = "rclr", MARGIN = 1)
crc_phe_list <- crc_phe_f |> dplyr::select(disease, study_name) |> base::split(f = crc_phe_f$study_name) 

res_wirbe <- .wilcox_all(crc_meta_clr, crc_phe_list$WirbelJ_2018)
res_yujun <- .wilcox_all(crc_meta_clr, crc_phe_list$YuJ_2015)
res_yachi <- .wilcox_all(crc_meta_clr, crc_phe_list$YachidaS_2019)

qdat <- data.frame(wirbe = res_wirbe$z_score, yujun = res_yujun$z_score, 
                   yachi = res_yachi$z_score, row.names = rownames(res_wirbe))
qdat[is.na(qdat)] <- 0
qdat_p <- data.frame(wirbe = res_wirbe$p.adjust, yujun = res_yujun$p.adjust, 
                   yachi = res_yachi$p.adjust, row.names = rownames(res_wirbe))
qdat_p[is.na(qdat_p)] <- 1

qdat_p_f <- qdat_p[apply(qdat_p, 1, function(x){any(x < 0.05)}), ]
qdat_p_f <- ifelse(qdat_p_f<0.05, 1, 0)
qdat_f <- qdat[rownames(qdat_p_f), ]
sigature <- sign(qdat_f)*qdat_p_f
qdat_p_num <- ifelse(sigature == 0, "", ifelse(sigature == 1, "+", "-") )

pheatmap::pheatmap(t(qdat_f), display_numbers = t(qdat_p_num), fontsize_number = 12)


save(res_wirbe, res_yachi, res_yujun, file = "crc_analysis_files/crc_clr_ra_compare.rds")

res_wirbe <- .wilcox_all(crc_meta_f, crc_phe_list$WirbelJ_2018)
res_yujun <- .wilcox_all(crc_meta_f, crc_phe_list$YuJ_2015)
res_yachi <- .wilcox_all(crc_meta_f, crc_phe_list$YachidaS_2019)

save(res_wirbe, res_yachi, res_yujun, file = "crc_analysis_files/crc_ra_compare.rds")


qdat <- data.frame(wirbe = res_wirbe$z_score, yujun = res_yujun$z_score, 
                   yachi = res_yachi$z_score, row.names = rownames(res_wirbe))
qdat[is.na(qdat)] <- 0
qdat_p <- data.frame(wirbe = res_wirbe$p.adjust, yujun = res_yujun$p.adjust, 
                   yachi = res_yachi$p.adjust, row.names = rownames(res_wirbe))
qdat_p[is.na(qdat_p)] <- 1

qdat_p_f <- qdat_p[apply(qdat_p, 1, function(x){any(x < 0.05)}), ]
qdat_p_f <- ifelse(qdat_p_f<0.05, 1, 0)
qdat_f <- qdat[rownames(qdat_p_f), ]
sigature <- sign(qdat_f)*qdat_p_f
qdat_p_num <- ifelse(sigature == 0, "", ifelse(sigature == 1, "+", "-") )

pheatmap::pheatmap(t(qdat_f), display_numbers = t(qdat_p_num), fontsize_number = 12)

```

### SIN based On spearman 

#### Compute the SIN 

```{r, warning=F, message=F}
library(igraph)
library(SpiecEasi)
library(interSNI)
library(metastat)
library(ggstatsplot)
library(birankr)

# here use clr-transformed 
datlist <- split(crc_meta_clr[rownames(crc_phe_f), ], crc_phe_f$study_name)
phelist <- crc_phe_list

# no-transformed data 

datlist2 <- split(as.data.frame(t(crc_meta[colnames(crc_meta_clr),rownames(crc_phe_f)]*10^7)), crc_phe_f$study_name)

# sweet 
sweet_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], ref = T)})
names(sweet_list) <- names(crc_phe_list)

# ssn
ssn_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "ssn",ref = T)})
names(ssn_list) <- names(crc_phe_list)

# lioness 
lioness_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "lioness",ref = T)})
names(lioness_list) <- names(crc_phe_list)

# iENA 
iENA_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "iENA",ref = F)})
names(iENA_list) <- names(crc_phe_list)

# csnet 
csnet_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "csnet",ref = F, boxsize = 0.2)})
names(csnet_list) <- names(crc_phe_list)

# sspg 
sspg_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "sspg", ref = F)})
names(sspg_list) <- names(crc_phe_list)

# save
save(sweet_list, ssn_list, lioness_list, iENA_list, csnet_list, sspg_list, file = "crc_analysis_files/Noref_spearman_ssn.rds")

```

### SIN based On sparcc 

```{r, warning=F, message=F}
# sweet 
sweet_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], ref = F, n_method = .sparccF)})
names(sweet_list_sparcc) <- names(crc_phe_list)
# ssn
ssn_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "ssn",ref = F, n_method = .sparccF)})
names(ssn_list_sparcc) <- names(crc_phe_list)
# lioness 
lioness_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "lioness",ref = F, n_method = .sparccF)})
names(lioness_list_sparcc) <- names(crc_phe_list)
# iENA 
iENA_list_sparcc <- iENA_list
csnet_list_sparcc <- csnet_list


save(sweet_list_sparcc, ssn_list_sparcc, lioness_list_sparcc, iENA_list_sparcc, csnet_list, sspg_list, file = "crc_analysis_files/Noref_sparCC_ssn.rds")

```


### SIN based On spearman and refernce data 

#### Compute the SIN 

```{r, warning=F, message=F}

crc_phe_ref <- crc_phe_f[crc_phe_f$disease == "healthy", ]
datlist_ref <- split(crc_meta_clr[rownames(crc_phe_ref), ], crc_phe_ref$study_name)

# sweet 
sweet_list_ref <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], ref = T, dat_ref = datlist_ref[[i]])})
names(sweet_list_ref) <- names(crc_phe_list)

# ssn
ssn_list_ref <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], method = "ssn",ref = T, dat_ref = datlist_ref[[i]])})
names(ssn_list_ref) <- names(crc_phe_list)

# lioness 
lioness_list_ref <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], method = "lioness",ref = T, dat_ref = datlist_ref[[i]])})
names(lioness_list_ref) <- names(crc_phe_list)

# iENA 
iENA_list_ref <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], method = "iENA",ref = T)})
names(iENA_list_ref) <- names(crc_phe_list)

# csnet 
csnet_list_ref <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "csnet",ref = T, boxsize = 0.2)})
names(csnet_list_ref) <- names(crc_phe_list)

# 
sspg_list_ref <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "sspg", ref = T)})
names(sspg_list_ref) <- names(crc_phe_list)


# save
save(sweet_list_ref, ssn_list_ref, lioness_list_ref, iENA_list_ref, csnet_list_ref, sspg_list_ref, file = "crc_analysis_files/ref_spearman_ssn.rds")

```

#### reference data

```{r, warning=F, message=F}

datlist2_ref <- split(as.data.frame(t(crc_meta[colnames(crc_meta_clr),rownames(crc_phe_ref)]*10^7)), crc_phe_ref$study_name)

# sweet 
sweet_list_ref_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "sweet", n_method = .sparccF, dat_ref = datlist2_ref[[i]])})

names(sweet_list_ref_sparcc) <- names(crc_phe_list)

# ssn
ssn_list_ref_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "ssn",ref = T, n_method = .sparccF, dat_ref = datlist2_ref[[i]])})
names(ssn_list_ref_sparcc) <- names(crc_phe_list)

# lioness 
lioness_list_ref_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "lioness",ref = T , n_method = .sparccF, dat_ref = datlist2_ref[[i]])})
names(lioness_list_ref_sparcc) <- names(crc_phe_list)

# iENA 
iENA_list_ref_sparcc <- iENA_list_ref

# csnet 
csnet_list_ref_sparcc <- csnet_list_ref

# sspg
sspg_list_ref_sparcc <- sspg_list_ref

# save
save(sweet_list_ref_sparcc, ssn_list_ref_sparcc, lioness_list_ref_sparcc, iENA_list_ref_sparcc, csnet_list_ref_sparcc, sspg_list_ref_sparcc, file = "crc_analysis_files/ref_sparcc_ssn.rds")

```

## Part3

### infant cohort 
```{r}
library(tidyverse)

infant_data <- readRDS("data/crc_infan.rds")
infant_phe <- infant_data$infant_meta[[1]]
infant_met <- infant_data$infant_meta[[2]]

infant_phe_f <- infant_phe[infant_phe$age_category != "adult", ]

library(gtsummary)
infant_phe_f |> select(infant_age, antibiotics_current_use, gender, number_reads, born_method) |> tbl_summary(by = born_method) |> add_p() |>
  add_overall() |>
  bold_labels()

```

```{r}
source("R/tinyFun.R")
library(RColorBrewer)
prevalences <- seq(0.05, 1, 0.05)
detections <- 10^seq(log10(1e-3), log10(.2), length = 10)

core_tax_plot(dat = t(infant_met), 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))


```

```{r}
id <- c("SID519128_ba", "SID515126_ba") 

sele_id <- infant_phe_f[infant_phe_f$subject_id %in% id, ] |> select(sample_id, infant_age, antibiotics_current_use, gender, number_reads, born_method)
sele_id

```

```{r}
library(rmeta)
infant_met_f <- filterPer(infant_met, row = 1, percent = 0.1) |> t() |> renorm()
infant_met_f <- na.omit(infant_met_f)
infant_met_f <- round(infant_met_f*10^5, 0)

infant_phe_f <- infant_phe_f[pmatch(rownames(infant_met_f), infant_phe_f$sample_id), ]

infant_C <- infant_phe_f[infant_phe_f$born_method == "c_section", ]
infant_C_ref4 <- infant_C[infant_C$infant_age ==4, ]
infant_C_ref <- infant_met_f[ infant_C_ref4$sample_id[!is.na(infant_C_ref4$sample_id)],] 
infant_C_met <- infant_met_f[infant_C$sample_id[!is.na(infant_C$sample_id)],]


infant_V <- infant_phe_f[infant_phe_f$born_method == "vaginal", ]
infant_V_ref4 <- infant_V[infant_V$infant_age ==4, ]
infant_V_ref <- infant_met_f[infant_V_ref4$sample_id[!is.na(infant_V_ref4$sample_id)],] 
infant_V_met <- infant_met_f[infant_V$sample_id[!is.na(infant_V$sample_id)],]

# SSN 
infant_c_sni <- ssn(matrix = infant_C_met, n_method = .sparccF, dat_ref = infant_C_ref)

infant_V_sni <- ssn(matrix = infant_V_met, n_method = .sparccF, dat_ref = infant_V_ref)
save(infant_c_sni, infant_V_sni, infant_phe_f, file = "data/infant_ssn.Rds")

```


