---
title: "CRC_inteSIN"
author: "Huahui Ren"
date: "2024-03-15"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduce 

Here is the test about the  classification power of CRC vs Control using sample-specific inference network methods.

## Content

- Part 1: Preprocessing 
- Part 2: Comparisions in biomarkers using relative abundance or degree 
- Part 3: Comparision in classifical performance 

## Part 1: 

```{r , message=FALSE, warning=F}
library(tidyverse)

crc_data <- readRDS("data/crc_infan.rds")
Yachi_meta <- crc_data$Yachidas_meta[[2]]  %>% as.data.frame() %>% mutate(id = rownames(.))
Yuj_meta <- crc_data$yuj_meta[[2]] %>% as.data.frame() %>% mutate(id = rownames(.))
Wir_meta <- crc_data$WirbelJ_meta[[2]]  %>% as.data.frame() %>% mutate(id = rownames(.))

crc_meta <- Yachi_meta %>%
  full_join(Yuj_meta, by = "id") %>%
  full_join(Wir_meta, by = "id") %>% column_to_rownames(var = "id")
crc_meta[is.na(crc_meta)] <- 0

library(rmeta)
crc_meta_f <- filterPer(crc_meta, row = 1, percent = 0.2) |> t() |> renorm()
crc_phe <- do.call("rbind", list(crc_data$Yachidas_meta[[1]], crc_data$yuj_meta[[1]], crc_data$WirbelJ_meta[[1]]))

crc_phe_f <- crc_phe |> filter(disease == "CRC" | disease == "healthy") 
rownames(crc_phe_f) <- crc_phe_f$sample_id
crc_meta_f <- crc_meta_f[rownames(crc_phe_f), ]

library(gtsummary)

crc_phe_f |> select(study_name, disease, age, gender, BMI) |> tbl_summary(by = study_name) |> add_p() |>
  add_overall() |>
  bold_labels()

```

## Part 2: 

### Overview of core species in 3 crc cohort 

*1.* WirbelJ_2018

```{r, message=F, error=F, warning=F}
library(RColorBrewer)
prevalences <- seq(0.05, 1, 0.05)
detections <- 10^seq(log10(1e-3), log10(.2), length = 10)

core_tax_plot <- function(dat, colors = gray(seq(0, 1, length=5)), prevalences, detections, min.prevalence, top){
  preva_tax <- names(which(apply(dat, 2, function(x){sum(x!=0)/length(x)}) > min.prevalence))
  median_tax <- names(sort(apply(dat, 2, median, na.rm = T), decreasing = T))[1:top] -> toptax
  retain_tax <- intersect(median_tax, preva_tax)
  dat_retain <- dat[, retain_tax]
  
  res_list <- lapply(detections, function(x){tmp <- dat_retain; tmp[dat_retain < x] <- 0; 
            pre <- apply(tmp,2,function(x){sum(x!=0)/length(x)});
            out <- data.frame(tax = retain_tax, prevalence = pre, cutoff = x)})
  
  qdat <- as.data.frame(do.call("rbind", res_list))
  qdat$tax <- factor(qdat$tax, levels = rev(retain_tax))
  qdat$cutoff <- paste0(round(qdat$cutoff*100, 3), "%")
  qdat$cutoff <- factor(qdat$cutoff, levels = paste0(round(detections*100, 3), "%"))
  
  p <- ggplot(qdat, aes(cutoff, tax, fill= prevalence)) + 
    geom_tile()+theme_bw()+theme(axis.text.x = element_text(angle = 90, hjust = 1))+
    scale_fill_gradient(low = colors[1], high = colors[5])
  
  return(p)
}

WirbelJ_2018_meta <- crc_meta_f[crc_phe_f$study_name == "WirbelJ_2018", ]

core_tax_plot(dat = WirbelJ_2018_meta, 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))

```

## 

*2.* YachidaS_2019

```{r}
YachidaS_2019_meta <- crc_meta_f[crc_phe_f$study_name == "YachidaS_2019", ]

core_tax_plot(dat = YachidaS_2019_meta , 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))
```

## 

*3.* YuJ_2015

```{r}
YuJ_2015_meta <- crc_meta_f[crc_phe_f$study_name == "YuJ_2015", ]

core_tax_plot(dat = YuJ_2015_meta , 
                          prevalences = prevalences, 
                          detections = detections, 
                          min.prevalence = 0.5, top = 30)+
  xlab("Detection Threshold (Relative Abundance (%))") +
  theme(axis.text.y = element_text(face="italic"))
```

## 

### Compare crc vs healthy using relative abundance  

```{r dplyr, warning=F, message=F}
library(coin)
source("R/tinyFun.R")
#first try to clr or alr transformation 
crc_meta_clr <- robCompositions::cenLR(zCompositions::cmultRepl(crc_meta_f))[[1]]
crc_phe_list <- crc_phe_f |> dplyr::select(disease, study_name) |> base::split(f = crc_phe_f$study_name) 

res_wirbe <- .wilcox_all(crc_meta_clr, crc_phe_list$WirbelJ_2018)
res_yujun <- .wilcox_all(crc_meta_clr, crc_phe_list$YuJ_2015)
res_yachi <- .wilcox_all(crc_meta_clr, crc_phe_list$YachidaS_2019)

qdat <- data.frame(wirbe = res_wirbe$z_score, yujun = res_yujun$z_score, 
                   yachi = res_yachi$z_score, row.names = rownames(res_wirbe))

qdat_p <- data.frame(wirbe = res_wirbe$p.adjust, yujun = res_yujun$p.adjust, 
                   yachi = res_yachi$p.adjust, row.names = rownames(res_wirbe))

qdat_p_f <- qdat_p[apply(qdat_p, 1, function(x){any(x < 0.05)}), ]
qdat_p_f <- ifelse(qdat_p_f<0.05, 1, 0)
qdat_f <- qdat[rownames(qdat_p_f), ]
sigature <- sign(qdat_f)*qdat_p_f
qdat_p_num <- ifelse(sigature == 0, "", ifelse(sigature == 1, "+", "-") )

pheatmap::pheatmap(t(qdat_f), display_numbers = t(qdat_p_num), fontsize_number = 12)

```

## SIN based On spearman 

### Compute the SIN 

```{r, warning=F, message=F}
library(igraph)
library(SpiecEasi)

datlist <- split(crc_meta_clr[rownames(crc_phe_f), ], crc_phe_f$study_name)
phelist <- crc_phe_list

datlist2 <- split(as.data.frame(t(crc_meta[colnames(crc_meta_clr),rownames(crc_phe_f)]*10^7)), crc_phe_f$study_name)


# sweet 
sweet_list <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], ref = T)})
names(sweet_list) <- names(crc_phe_list)
# ssn
ssn_list <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], method = "ssn",ref = T)})
names(ssn_list) <- names(crc_phe_list)
# lioness 
lioness_list <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], method = "lioness",ref = T)})
names(lioness_list) <- names(crc_phe_list)
# iENA 
iENA_list <- lapply(1:3, function(i){intnet_com(dat = datlist[[i]], phe = phelist[[i]], method = "iENA",ref = T)})
names(iENA_list) <- names(crc_phe_list)
# cenet 
# 这里使用
csnet_list <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "csnet",ref = F, boxsize = 0.2)})
names(csnet_list) <- names(crc_phe_list)

View(cbind(res_wirbe_degree[rownames(res_yachi), ], res_yachi))

```

## SIN based On sparcc 

```{r, warning=F, message=F}
# sweet 
sweet_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], ref = F, n_method = .sparccF)})
names(sweet_list_sparcc) <- names(crc_phe_list)
# ssn
ssn_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "ssn",ref = F, n_method = .sparccF)})
names(ssn_list_sparcc) <- names(crc_phe_list)
# lioness 
lioness_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "lioness",ref = F, n_method = .sparccF)})
names(lioness_list_sparcc) <- names(crc_phe_list)
# iENA 
iENA_list_sparcc <- lapply(1:3, function(i){intnet_com(dat = datlist2[[i]], phe = phelist[[i]], method = "iENA",ref = F)})
names(iENA_list_sparcc) <- names(crc_phe_list)


View(cbind(res_wirbe_degree[rownames(res_yachi), ], res_yachi))

```




```{r}
intnet_com  <- function(dat, phe, method = "sweet", n_method = .spearmanF, ref=F, ...){
  # filter 
  
  if(method == "sweet"){
    sweet_res <- Sweet(dat = dat, s_method = .spearmanF, n_method = n_method)
    netres <- .net_stat_com(sweet_res)
  }else if(method == "ssn"){
    ssn_res <- ssn(matrix = dat,  n_method = n_method, transF = F)
    netres <- .net_stat_com(ssn_res)  
  }else if(method == "lioness"){
    lioness_res <- lioness(matrix = dat,  n_method = n_method, transF = F, alpha = 0.05)
    netres <- .net_stat_com(lioness_res) 
  }else if(method == "iENA"){
    if(ref){
      h_id <- rownames(phe[phe$disease == "healthy", ])
      healthy <- dat[h_id, ]
      ref_mean <- apply(healthy, 2, mean)
      ref_var <- apply(healthy, 2, var)
      iENA_res <- iENA_sPCC(dat  = dat, transF = F, ref = T, refmean = ref_mean, refvar = ref_var)
    }else{
      iENA_res <- iENA_sPCC(dat  = dat, transF = F, ref = F)
    }
      netres <- .net_stat_com(iENA_res) 
  }else if(method == "csnet"){
    if(ref){
      h_id <- rownames(phe[phe$disease == "healthy", ])
      healthy <- dat[h_id, ]
      nh_id <- rownames(phe[phe$disease != "healthy", ])
      nhealthy <- dat[nh_id, ]
      csnet_res_h <-  csnet(dat  =  t(healthy))
      csnet_res_nh <-  csnet(dat  =  t(nhealthy))
      netres_h <- .net_stat_com(csnet_res_h)
      netres_nh <- .net_stat_com(csnet_res_nh)
      netres <- list(edge_number = rbind(netres_h$edge_number, netres_nh$edge_number),
                     degree = cbind(netres_h$degree, netres_nh$degree),
                     betweeness = cbind(netres_h$betweeness, netres_nh$betweeness),
                     closeness = cbind(netres_h$closeness, netres_nh$closeness))
    }else{
      csnet_res <- csnet(dat  =  t(dat), ...)
      netres <- .net_stat_com(csnet_res)
    }
  }
  # degree 
  degree_matrix <- t(netres$degree)
  res_degree <- .wilcox_all(degree_matrix, phe)
  # betweeness 
  betweeness_matrix <- t(netres$betweeness)
  res_betweeness <- .wilcox_all(betweeness_matrix, phe)  
  # closeness 
  closeness_matrix <- t(netres$closeness)
  closeness_matrix[is.na(closeness_matrix)] <- 0
  res_closeness <- .wilcox_all(closeness_matrix, phe)   
  # edge compare 
  id <- intersect(rownames(netres$edge_number), rownames(phe))
  edge_dat <- data.frame(netres$edge_number[id, ], phe[id, ])
  qdat <- melt(edge_dat)
  com_plot <- grouped_ggbetweenstats(data = qdat, x = disease, y = value, grouping.var = variable, bf.message = F)
  # output 
  outlist <- list(res_degree, res_betweeness, res_closeness, com_plot)
  names(outlist) <- c("degree", "betweeness", "closeness", "edge")
  return(outlist)
}
  
```





