---
title: "Microbiome test"
output: html_notebook
---

## 1. load data

```{r, warning=F}
library(data.table)
dat <- fread("data/gyy.metadata.tsv", data.table = F) # 
micro <- fread("data/gyy.mp.tsv", data.table = F)
rownames(dat) <- dat[,1]
metadata <- dat[,-1]
colnames(micro) <- gsub(".metaphlan4.abundance.profile", "", colnames(micro))
rownames(micro) <- micro$clade_name
micro <- micro[,-c(1:2)]
```

### 1.1 preprocess

```{r fig.height=4, fig.width=8, warning=FALSE, paged.print=TRUE}
# preprocess 
datalist <- .splitMetaphlan(data = micro, prefix = "drug")
genus <- datalist$drug_genus
genus_filter <- genus[apply(genus, 1, function(x){sum(x>0.01)>20}), ] # filter rare genus
genus_norm <- t(genus_filter)/colSums(genus_filter) # renorm genus 

# plot the top10 genus composition 
library(RColorBrewer)
.topTax(t(genus_norm), K = 10, tax_colour = brewer.pal(n = 12, name = "Paired"))
```

### 1.2 SIN generate using Sweet

```{r}
# first to test sweet 
psudo_count <- round(genus_norm*10^7,0)
sweet_res <- Sweet(dat = psudo_count, s_method = .spearmanF, n_method = .sparccF)
save(sweet_res, file = "data/gyy_sweet.Rdata")
```


### 1.3 SIN analysis

#### 1.3.1 Comparison of SIN degree between different group 

```{r}
load("data/gyy_sweet.Rdata")
library(igraph)
library(SpiecEasi)
source("R/tinyFun.R")
cutoff <- -qnorm(0.01, mean = 0, sd = 1)
outlist <- lapply(sweet_res, function(x){x[abs(x) < cutoff] <- 0; x <- sign(x); as(x, "dgCMatrix")})
## degree comparision  
netres <- .net_stat_com(outlist)
degree_matrix <- netres$degree
rownames(degree_matrix) <- colnames(psudo_count)
```

Comparison of SIN degree between pre-treatment and post-treatment in **acarbose**

```{r}
##----------------- acar -------------------------------------- ## 
acar <- metadata[metadata$Drug == "Acarbose", ]
library(coin)
library(rcompanion)

acar_degree_res <- .mywilcox_2t(datamatrix = t(degree_matrix), configdata = acar, group = "Time",  ID = "PID", Time = "Time")
library(DT)
datatable(acar_degree_res, caption = "Comparision of degree between pre-treatment and post-treatment in acarbose group") %>% formatRound(colnames(acar_degree_res)[1:10],3)
```

Comparison of SIN degree between pre-treatment and post-treatment in **glipizide**

```{r}
##---------------- glip --------------------------------------- ##
Glip <- metadata[metadata$Drug == "Glipizide", ]
glip_degree_res <- .mywilcox_2t(datamatrix = t(degree_matrix), configdata = Glip, group = "Time",  ID = "PID", Time = "Time")

datatable(glip_degree_res, caption = "Comparision of degree between pre-treatment and post-treatment in Glipizide group") %>% formatRound(colnames(glip_degree_res)[1:10],3)
```

#### 1.3.2 Comparison of SIN degree matrix between different group 

acarbose group 

```{r}
## permanova 
acar_degree <- t(degree_matrix[,rownames(acar)])
acar_perm <- adonis2(acar_degree~Time, data = acar, method = "jaccard")
datatable(acar_perm)
```

glipizide group 

```{r}
glip_degree <- t(degree_matrix[,rownames(Glip)])
glip_perm <- adonis2(glip_degree~Time, data = Glip, method = "jaccard")
datatable(glip_perm)
```


#### 1.3.2 Comparison of edge between different group 

acarbose group 

```{r}
## gene_pair comparision 
library(reshape2)
data_pair <- melt(lapply(outlist, function(x){tmp <- as.matrix(x); diag(tmp) <- NA; tmp[lower.tri(tmp)] <- NA;tmp}))

data_pair <- data_pair[!is.na(data_pair[,3]),] # remove the loop and repeat edga 
colnames(data_pair) <- c("feature1", "feature2", "edga", "sample")
data_pair$pair <- paste0(data_pair$feature1, "_", data_pair$feature2)
pair_data <- acast(data_pair, sample ~ pair, value.var = "edga")
pair_data_filter <- pair_data[,apply(pair_data, 2, function(x){sum(x!=0)>20})]

## acar 
pair_acar <- pair_data_filter[rownames(acar), ]
acar_pair_res <- .mywilcox_2t(datamatrix = pair_acar, configdata = acar, group = "Time",  ID = "PID", Time = "Time")
bac_name <- colnames(psudo_count)
acar_pair_res$feature1 <- unlist(lapply(rownames(acar_pair_res), function(x){index <- as.numeric(unlist(strsplit(x, split = "_"))); bac_name[index[1]]}))
acar_pair_res$feature2 <- unlist(lapply(rownames(acar_pair_res), function(x){index <- as.numeric(unlist(strsplit(x, split = "_"))); bac_name[index[2]]}))

acar_pair_res_sig <- acar_pair_res[acar_pair_res$sign_p.value < 0.01, ]
pheatmap(pair_acar[, rownames(acar_pair_res_sig)], annotation_row = acar[,'Time',drop=F],show_rownames = F)
```

glipizide group 

```{r}
## glip 
pair_glip <- pair_data_filter[rownames(Glip), ]
glip_pair_res <- .mywilcox_2t(datamatrix = pair_glip, configdata = Glip, group = "Time",  ID = "PID", Time = "Time")
glip_pair_res$feature1 <- unlist(lapply(rownames(glip_pair_res), function(x){index <- as.numeric(unlist(strsplit(x, split = "_"))); bac_name[index[1]]}))
glip_pair_res$feature2 <- unlist(lapply(rownames(glip_pair_res), function(x){index <- as.numeric(unlist(strsplit(x, split = "_"))); bac_name[index[2]]}))

glip_pair_res_sig <- glip_pair_res[glip_pair_res$sign_p.value < 0.01, ]
pheatmap(pair_glip[, rownames(glip_pair_res_sig)], annotation_row = Glip[,'Time',drop=F],show_rownames = F)
```

